---
title: "Full Stack with R"
description: |
  How to do 'Full Stack Web Development' with R
author: Kyle Tolliver
date: 03-26-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
pacman::p_load(tidyverse, pacman, knitr, rmarkdown, pander, glue)
opts_chunk$set(results = 'asis', echo = FALSE, message = FALSE, warning = FALSE) ## Chunk Displays
```

```{r functions}
info <- function(url){pander(glue("See [{url}]({url}) for more information"))}

commands <- function(package){
  pander(glue('
  Basic Commands

    install.package("{package}")
    library({package})
'))
}
```


This page is a tutorial of what packages and tools could be used to mimic Full Stack Web Development with R. 

## Tidyverse

`r commands("tidyverse")`

Tidyverse is a collection of useful packages for Data Science. I believe its no mater what you use R for.

Packages include:

* ggplot2 - System for declaratively creating graphics, based on The Grammar of Graphics
* dplyr - Provides a grammar of data manipulation, providing a consistent set of verbs that solve the most common data manipulation challenges
* tidyr - Provides a set of functions that help you get to tidy data
  - Tidy data is data with a consistent form: in brief, every variable goes in a column, and every column is a variable
* readr - Provides a fast and friendly way to read rectangular data (like csv, tsv, and fwf)
  - Similar packages include:
    + DBI
    + haven
    + httr
    + readxl
    + rvest
    + jsonlite
    + xml2
* purrr - Enhances R’s functional programming (FP) toolkit by providing a complete and consistent set of tools for working with functions and vectors
  - Also includes 
    + magrittr - Provides the pipe, %>% used throughout the tidyverse
    + glue - Provides an alternative to paste() that makes it easier to combine data and strings
* tibble - Modern re-imagining of the data frame, keeping what time has proven to be effective, and throwing out what it has not
* stringr - Provides a cohesive set of functions designed to make working with strings as easy as possible
* forcats - Provides a suite of useful tools that solve common problems with factors

`r info("https://www.tidyverse.org/")`

## Pacman

`r commands("pacman")`
    
The pacman package is an R package management tool that combines the functionality of base library related functions into intuitively named functions.

```{r pacman}
pacman <- readr::read_csv('./pacman.csv')
pander::pander(pacman)
```

`r info("https://rdocumentation.org/packages/pacman/versions/0.5.1")` 

## Googlesheets4

`r commands("googlesheets4")`
    
googlesheets4 will, by default, help you interact with Sheets as an authenticated Google user. 

Use `gs4_deauth()` to indicate there is no need for a token

Similar Packages include:

* googlesheets - Package that googlesheets4 replaces
* googledrive - Provides a fully-featured interface to the Google Drive API 
* readxl - Tidyverse package for reading Excel files
* readr - tidyverse package for reading delimited files

`r info("https://googlesheets4.tidyverse.org/")`

## RMarkdown and Knitr

`r commands("rmarkdown")`
`r commands("knitr")`

Allows for flexiblity of files and functions when converting to html.

`r info("https://rmarkdown.rstudio.com/, https://www.rdocumentation.org/packages/knitr/versions/1.31")`

## Pander and Glue

### Pander

`r commands("pander")`

Prints an R object in Pandoc's markdown. Used to beautify output in R. 

`r info("https://www.rdocumentation.org/packages/pander/versions/0.6.3")`

### Glue

`r commands("glue")`

Glue offers interpreted string literals that are small, fast, and dependency-free. Glue does this by embedding R expressions in curly braces which are then evaluated and inserted into the argument string.

Basic Functions

    glue()
    glue_data()
    glue_sql()

`r info("https://glue.tidyverse.org/")`

### Pander and Glue together

Allow for reuse of print statements. 

For example `info <- function(url){pander(glue("See [{url}]({url}) for more information"))` was used in this page and I just inserted the url into the function call. 

## SQL Packages

SQL - database query language -- a language designed specifically for interacting with a database

* sqldf
* DBI
* RODBC
* dbConnect
* RSQLite
* RMySQL
* RPostgreSQL

These packages allow you to run SQL commands in R.

`r info("https://dept.stat.lsa.umich.edu/~jerrick/courses/stat701/notes/sql.html")`

## Jsonlite and V8

The jsonlite package is a JSON parser/generator optimized for the web. 
`r info("https://cran.r-project.org/web/packages/jsonlite/vignettes/json-aaquickstart.html")`

V8 is Google’s open source, high performance JavaScript engine. V8 R package allows you implement JS in R. 
`r info("https://cran.r-project.org/web/packages/V8/vignettes/v8_intro.html")`

## ReactR

reactR provides a set of convenience functions for using React in R with htmlwidget constructor templates and local JavaScript dependencies. The React ecosystem is rich with components that can enhance R web and Shiny apps.

Learn ReactJS at https://www.w3schools.com/react/default.asp

`r info("https://react-r.github.io/reactR/")`

## Shiny

Shiny is an R package that makes it easy to build interactive web apps straight from R. You can host standalone apps on a webpage or embed them in R Markdown documents or build dashboards. 

`r info("https://shiny.rstudio.com/")`

## Other

* leaflet
* fs
* curl
* distill
* devtools
* bookdown
* pagedown